Quasicharacters and the function-sheaf dictionary

### 2014-09-10 crelle-dictionary.tex
Messing with bibliography

%%%%%%%%%%%%%%%%% PACKAGES %%%%%%%%%%%%%%%%%
% Encodings
%\usepackage[utf8]{inputenc}
%\usepackage[T1]{fontenc}
% AMS packages
%\usepackage{amsmath, amsthm, amssymb}
\usepackage{todonotes}%, showkeys}
\usepackage{amsrefs}
\BibSpec{lecture}{%
  +{}{\PrintAuthors} {author}
  +{,}{ \textit} {title}
  +{}{ \parenthesize} {date}
  +{,}{ } {note}
  +{,}{ } {location}
  +{.}{ } {transition}
}
\BibSpec{preprint}{%
  +{}{\PrintAuthors} {author}
  +{,}{ \textit} {title}
  +{,}{ preprint } {date}
  +{,}{ } {note}
  +{.}{ } {transition}
}
% Fonts
\usepackage{mathrsfs}
%\usepackage{yfonts}
%Geometry
%\usepackage{geometry}
% Enumitem
\usepackage{enumitem}
% TikZ
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,calc,matrix}
\usepackage{tikz-cd}
% Hyperrefs
\usepackage{hyperref}

%%%%%%%%%%%%%%% THEOREM STYLES %%%%%%%%%%%%%%%
\theoremstyle{plain}
      \newtheorem{theorem}{Theorem}[section]
      \newtheorem*{theorem*}{Theorem}
      \newtheorem{proposition}[theorem]{Proposition}
      \newtheorem{lemma}[theorem]{Lemma}
      \newtheorem{corollary}[theorem]{Corollary}

@@

%\newcommand{\Zl}{\mathbb{\bar Z}_\ell}

@@

%\newcommand{\Zlx}{\mathbb{\bar Z}_\ell^\times}

@@

%\DeclareMathOperator{\W}{W}

@@

\begin{abstract}
We introduce the category of quasicharacter sheaves on smooth commutative group schemes $G$ over finite fields $k$ and expand the scope of the function-sheaf dictionary from connected commutative algebraic groups to this setting.
We show that the group of isomorphism classes of quasicharacter sheaves is an extension of the group of characters of $G(k)$ by a cohomology group determined by the component group scheme of $G$.
As an application, we find a category of sheaves for which isomorphism classes correspond to quasi\-characters in the case of algebraic tori, abelian varieties and unipotent wound groups over local fields. 
In this geometrization of quasicharacters, our sheaves are local systems on
Greenberg transforms of locally finite type N\'eron models.
\end{abstract}

@@

Motivated by an application to the geometrization and categorification of quasicharacters of $p$-adic tori,
this paper extends Deligne's function-sheaf dictionary to smooth commutative group schemes over finite fields. 
\footnote{Mathematics Subject Classification 2010: 14F05 (primary), 14L15 (secondary), 22E50 (tertiary)}

@@

By any measure, there are more quasicharacter sheaves for $T$ than quasicharacters of $T(K)$.  
In this regard, we are reminded of the work of David Vogan \cite{vogan:93a}, in which he finds a geometrization of complete Langlands parameters for $p$-adic groups 
and, in the process, is led to study the representations of all the pure rational forms of the $p$-adic group, simultaneously.
A similar phenomenon appears in recent work by Joseph Bernstein in which his geometric Ansatz leads to the study of certain sheaves on the stacky classifying space of the $p$-adic group, resulting in a category which appears to be tied to the representations of all the pure rational forms of the $p$-adic group \cite{bernstein:vogan_conference}.
Indeed, Bernstein has suggested to us that our category of quasicharacters for $T$ may be tied to quasicharacters of all the pure rational forms of $T$.

@@

\begin{proof}
Choose any $\Fq$-rational point $g$ on $G$ and let $\bg$ be the geometric point on $G$ lying above $g$.
Recall that the \emph{Weil group} of $G$, which we will denote by $\Weil{}(G,\bg)$, is a subgroup of the \'etale
fundamental group defined by the following diagram:
\[
 \begin{tikzcd}
 1 \rar & \ar[equal]{d} \pi_1(\bG, \bg) \rar & \Weil{}(G,\bg) \rar \dar[hook] & \Weil{} \rar \dar[hook] & 1 \\
 1 \rar &  \pi_1(\bG, \bg) \rar & \pi_1(G,\bg) \rar & \Gal(\bFq/\Fq) \rar & 1.
 \end{tikzcd}
\]
The $\Fq$-rational point $g$ under the geometric point $\bg$ determines a splitting
$\Weil{}\to \Weil{}(G,\bg)$ of $\Weil{}(G,\bg)\to \Weil{}$.
%
  Since $G$ is connected, the geometric point $\bg$ determines
  an equivalence between the category of $\ell$-adic Weil local systems on $G$ and
  $\ell$-adic representations of $\Weil{}(G,\bg)$ \cite{deligne:80a}*{1.1.12}.

  Now let $(\gqcs{L},\mu,\phi)$ be a quasicharacter sheaf on $G$
  and let $\lambda : \Weil{}(G, \bg) \to \EEx$ be the character determined by $(\gqcs{L},\phi)$.
  Composing with the splitting $\Weil{} \to \Weil{}(G,\bg)$ yields an $\ell$-adic character
  $\lambda_g : \Weil{} \to \EEx$, which is the same as the Trace of Frobenius defined in Section~\ref{ssec:Frob}, for every $\Fq$ rational point $g$ on $G$:
  $
  \lambda_g(\Frob{}) =  \trFrob{\qcs{L}}(g).
  $
%
  On the other hand, we have already seen that $\trFrob{\qcs{L}} : G(\Fq) \to \EEx$
  is a group homomorphism. 
  Since $G$ is an algebraic group over $\Fq$, $G(\Fq)$ is finite.
  Therefore $\trFrob{\qcs{L}}(g) = \lambda_g(\Frob{})$ is a root of unity
  for every $g\in G(\Fq)$.  Since $\Weil{}$ is generated by
  $\Frob{}$ and $\lambda_g : \Weil{} \to \EEx$ is
  a character, it follows that the image of $\lambda_g$ is a finite group.
  Thus, $\lambda_g$ extends to an $\ell$-adic character of $\Gal(\bFq/\Fq)$,
  which we will also denote $\lambda_g$.
%
  We may now lift the $\ell$-adic character $\lambda_g : \Gal(\bFq/\Fq) \to \EEx$
  to an $\ell$-adic character $\pi_1(G,\bg) \to \EEx$ using the canonical topological group homomorphism
  $\pi_1(G,\bg) \to \Gal(\bFq/\Fq)$. 
 % 
  The $\Fq$ rational point $g$ also
  determines an equivalence between the category of $\ell$-adic
  representations of $\pi_1(G,\bg)$ and $\ell$-adic local systems on $G$. Let
  $\qcs{E}$ be a local system on $G$ in the isomorphism class
  determined by this $\ell$-adic character of $\pi_1(G,\bg)$.
  Then $b_G^*\qcs{E} \iso \gqcs{L}$.

@@

%-------------------------------------------------------------------------%
% BIBLIOGRAPHY                                                            %
% References should be collected at the end of the paper and numbered in  %
% alphabetical order of the authors' names. Titles of journals should be  %
% abbreviated as in Mathematical Reviews. Please do not use small caps    %
% for the names of the authors. The preferred style is shown in the       %
% examples below.                                                         %
% Refer to the bibliography entries using \cite commands.                 %
% Using BibTeX, you simply have to add the names of your databases as     %
%   \bibliography{mydatabase}                                             %
% The file crelle.bst then ensures the correct style.                     %
%-------------------------------------------------------------------------%
%\bibliographystyle{amsalpha}
%\bibliography{../bibliography/Biblio}

>>>>

%%%%%%%%%%%%%%%%% PACKAGES %%%%%%%%%%%%%%%%%
% Encodings
%\usepackage[utf8]{inputenc}
%\usepackage[T1]{fontenc}
% AMS packages
%\usepackage{amsmath, amsthm, amssymb}
\usepackage{todonotes}%, showkeys}
%\usepackage[numeric,initials]{amsrefs}
\usepackage{crellerefs}

% Fonts
\usepackage{mathrsfs, yfonts}
%Geometry
%\usepackage{geometry}
% Enumitem
\usepackage{enumitem}
% TikZ
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,calc,matrix}
\usepackage{tikz-cd}
% Hyperrefs
\usepackage{hyperref}

%%%%%%%%%%%%%%% COUNTER %%%%%%%%%%%%%%%
%\setcounter{section}{-1}
%%%%%%%%%%%%%%% THEOREM STYLES %%%%%%%%%%%%%%%
\theoremstyle{plain}
      \newtheorem{theorem}{Theorem}[section]
      \newtheorem*{theorem*}{Theorem}
      \newtheorem{proposition}[theorem]{Proposition}
      \newtheorem{lemma}[theorem]{Lemma}
      \newtheorem{corollary}[theorem]{Corollary}

@@

\newcommand{\Zl}{\mathbb{\bar Z}_\ell}

@@

\newcommand{\Zlx}{\mathbb{\bar Z}_\ell^\times}

@@

\DeclareMathOperator{\W}{W}

@@

\begin{abstract}
We introduce the category of quasicharacter sheaves on smooth commutative group schemes $G$ over finite fields $k$ and expand the scope of the function-sheaf dictionary from connected commutative algebraic groups to this setting.
We show that the group of isomorphism classes of quasicharacter sheaves is an extension of the group of characters of $G(k)$ by a cohomology group determined by the component group scheme of $G$.
As an application, we find a category of sheaves for which isomorphism classes correspond to quasi\-characters in the case of algebraic tori, abelian varieties and unipotent wound groups over local fields. 
In this geometrization of quasicharacters, our sheaves are local systems on
Greenberg transforms of locally finite type N\'eron models.
Mathematics Subject Classification 2010: 14F05 (primary), 14L15 (secondary)
\end{abstract}

@@

Motivated by an application to the geometrization and categorification of quasicharacters of $p$-adic tori,
this paper extends Deligne's function-sheaf dictionary to smooth commutative group schemes over finite fields.

@@

By any measure, there are more quasicharacter sheaves for $T$ than quasicharacters of $T(K)$.  
In this regard, we are reminded of the work of David Vogan \cite{vogan:93a}, in which he finds a geometrization of complete Langlands parameters for $p$-adic groups 
and, in the process, is led to study the representations of all the pure rational forms of the $p$-adic group, simultaneously.
A similar phenomenon appears in recent work by Joseph Bernstein in which his geometric Ansatz leads to the study of certain sheaves on the stacky classifying space of the $p$-adic group, resulting in a category which appears to be tied to the representations of all the pure rational forms of the $p$-adic group %\cite{bernstein:vogan_conference}.
Indeed, Bernstein has suggested to us that our category of quasicharacters for $T$ may be tied to quasicharacters of all the pure rational forms of $T$.

@@

\begin{proof}
Choose any $\Fq$-rational point $g$ on $G$ and let $\bg$ be the geometric point on $G$ lying above $g$.
Recall that the \emph{Weil group} of $G$, which we will denote by $\W(G,\bg)$, is a subgroup of the \'etale
fundamental group defined by the following diagram:
\[
 \begin{tikzcd}
 1 \rar & \ar[equal]{d} \pi_1(\bG, \bg) \rar & \W(G,\bg) \rar \dar[hook] & \Weil{} \rar \dar[hook] & 1 \\
 1 \rar &  \pi_1(\bG, \bg) \rar & \pi_1(G,\bg) \rar & \Gal(\bFq/\Fq) \rar & 1.
 \end{tikzcd}
\]
The $\Fq$-rational point $g$ under the geometric point $\bg$ determines a splitting
$\Weil{}\to \W(G,\bg)$ of $\W(G,\bg)\to \Weil{}$.
%
  Since $G$ is connected, the geometric point $\bg$ determines
  an equivalence between the category of $\ell$-adic Weil local systems on $G$ and
  $\ell$-adic representations of $\W(G,\bg)$ \cite{deligne:80a}*{1.1.12}.

  Now let $(\gqcs{L},\mu,\phi)$ be a quasicharacter sheaf on $G$
  and let $\lambda : \W(G, \bg) \to \EEx$ be the character determined by $(\gqcs{L},\phi)$.
  Composing with the splitting $\Weil{} \to \W(G,\bg)$ yields an $\ell$-adic character
  $\lambda_g : \Weil{} \to \EEx$, which is the same as the Trace of Frobenius defined in Section~\ref{ssec:Frob}, for every $\Fq$ rational point $g$ on $G$:
  $
  \lambda_g(\Frob{}) =  \trFrob{\qcs{L}}(g).
  $
%
  On the other hand, we have already seen that $\trFrob{\qcs{L}} : G(\Fq) \to \EEx$
  is a group homomorphism. 
  Since $G$ is an algebraic group over $\Fq$, $G(\Fq)$ is finite.
  Therefore $\trFrob{\qcs{L}}(g) = \lambda_g(\Frob{})$ is a root of unity
  for every $g\in G(\Fq)$.  Since $\Weil{}$ is generated by
  $\Frob{}$ and $\lambda_g : \Weil{} \to \EEx$ is
  a character, it follows that the image of $\lambda_g$ is a finite group.
  Thus, $\lambda_g$ extends to an $\ell$-adic character of $\Gal(\bFq/\Fq)$,
  which we will also denote $\lambda_g$.
%
  We may now lift the $\ell$-adic character $\lambda_g : \Gal(\bFq/\Fq) \to \EEx$
  to an $\ell$-adic character $\pi_1(G,\bg) \to \EEx$ using the canonical topological group homomorphism
  $\pi_1(G,\bg) \to \Gal(\bFq/\Fq)$. 
 % 
  The $\Fq$ rational point $g$ also
  determines an equivalence between the category of $\ell$-adic
  representations of $\pi_1(G,\bg)$ and $\ell$-adic local systems on $G$. Let
  $\qcs{E}$ be a local system on $G$ in the isomorphism class
  determined by this $\ell$-adic character of $\pi_1(G,\bg)$.
  Then $b_G^*\qcs{E} \iso \gqcs{L}$.

@@

%-------------------------------------------------------------------------%
% BIBLIOGRAPHY                                                            %
% References should be collected at the end of the paper and numbered in  %
% alphabetical order of the authors' names. Titles of journals should be  %
% abbreviated as in Mathematical Reviews. Please do not use small caps    %
% for the names of the authors. The preferred style is shown in the       %
% examples below.                                                         %
% Refer to the bibliography entries using \cite commands.                 %
% Using BibTeX, you simply have to add the names of your databases as     %
%   \bibliography{mydatabase}                                             %
% The file crelle.bst then ensures the correct style.                     %
%-------------------------------------------------------------------------%
\bibliographystyle{crelle}
\bibliography{bibliography/Biblio}

% italicized and
% Bernstein lecture
% Etingof url, preprint note
% Italicized author in D-G appendix
% EGA ordering
% Ravi notes
% Vogan LLC
